\PassOptionsToPackage{svgnames,dvipsnames}{xcolor}

\documentclass[11pt]{cmuthesis}

\usepackage[Lenny]{fncychap}
\ChNameVar{\Large}

\input{packages}
\input{macros}

%\draftstamp{\today}{DRAFT}

\begin{document}
\frontmatter

\pagestyle{empty}

\title{{\bf Planning and Execution using Inaccurate Models with
    Provable Guarantees on Task Completeness}}
\author{Anirudh Vemula}
\date{February 2 2022}
\Year{2022}
\trnumber{CMU-RI-TR-22-05}

\committee{
  Maxim Likhachev, Co-Chair \\
  J. Andrew Bagnell, Co-Chair \\
  Oliver Kroemer \\
  Leslie Pack Kaelbling, \textit{Massachusetts Institute of Technology}
%   \begin{tabular}{cc}
%     & \\
% Maxim Likhachev, Co-Chair & CMU\\
% J. Andrew Bagnell, Co-Chair & CMU\\
%     Oliver Kroemer & CMU \\
%     Leslie Pack Kaelbling & MIT
% \end{tabular}
}

\support{}
\disclaimer{}

\keywords{Robotics, Planning, Reinforcement Learning, Numerical Optimization}

\maketitle

\begin{dedication}
  % For my brother, who put the fear of mediocrity in me \\
  For my mother, who sacrificed her own dreams to help me pursue
  mine
  % For my love, who believed in me and brought me back to surface
  % when I sank to the depths
\end{dedication}

\begin{abstract}
  Modern planning methods are effective in computing feasible and
optimal plans for robotic tasks when given access to accurate
dynamical models.
However, robots operating in the real
world often face situations that cannot be modeled perfectly before
execution.
Thus, we only have access to simplified but potentially inaccurate 
models.
This imperfect modeling can lead to highly suboptimal plans
or even the inability to reach the goal during execution.
Existing approaches present a learning-based solution where real-world
experience is used to learn a complex dynamical model that is
subsequently used for planning. However, this requires a prohibitively
large amount of experience over the entire state space, and can
be wasteful if we are interested in completing the task and not in
modeling the dynamics accurately. Furthermore, real robots often have
operating constraints and cannot spend hours acquiring experience to
learn dynamics.
This thesis argues that by updating the behavior of the planner and
not the dynamics of the model, we can
leverage simplified and potentially inaccurate models and
significantly reduce the amount of real-world experience needed to
provably guarantee that the robot completes the task.

We support this argument from an algorithmic perspective
by presenting two novel algorithms.
The first algorithm \cmax{} guarantees that the robot reaches the
goal using the inaccurate model without any resets. This
is achieved by biasing the planner away from transitions whose
dynamics are discovered to be inaccurately modeled during online
execution. However, \cmax{} requires strong assumptions on the
accuracy of the model used for planning and fails to improve the
quality of solution over repetitions of the same task. The second
algorithm \cmaxpp{} leverages real-world experience to improve the
quality of resulting plans over successive repetitions of a robotic
task. \cmaxpp{} achieves this by integrating model-free learning using
acquired experience with model-based planning using the potentially
inaccurate model. As a consequence of this in addition to completeness, \cmaxpp{} also
guarantees asymptotic convergence to the optimal path cost
as the number of repetitions increases under relaxed
assumptions. Crucially, both algorithms do not require any updates to
the dynamics of the model unlike any existing method for planning
using inaccurate models.

From a theoretical perspective,
this thesis presents a
performance analysis for methods
that leverage inaccurate models in optimal control of linearized systems with
quadratic costs. Our analysis shows that naively using inaccurate
models can lead to large suboptimality gaps when modeling errors are
significant, while updating the behavior of the planner during
execution, like \cmax{} and \cmaxpp{}, can substantially reduce the
suboptimality gap. The thesis concludes by exploring the paradigm of
updating the dynamics of the model and presents an algorithm \taml{} that
directly optimizes task performance rather than prediction error. We
show that in the online setting where the 
robot does not have access to any resets and collects data as it
executes, \taml{} outperforms prior 
works that either optimize a maximum likelihood objective or rely on
an offline collected dataset with good coverage.

\end{abstract}

\newgeometry{left=0.5in,right=0.5in,top=1in,bottom=1.4in}
\begin{acknowledgments}
  I believe PhD is all about the journey and not about the destination. I was
  extremely fortunate to have a
  wonderful journey
  with amazing people who have been my collaborators, confidantes,
  friends, mentors, and family. I will try my best to thank each and every one of them
  but alas, my memory often fails me. So if I forget someone below, know that I am
  lucky to have known them and they made my journey all the more fun.
  % While I leave with a doctorate in my hand, I cherish my
  % interactions with them the most and will always look back fondly on these
  % memories.

  % Advisors - Max, Drew
  First and foremost, I would like to thank my advisors Max and Drew. I always
  wanted to work with both of them, and forging this joint
  collaboration remains as the most important decision I have made in
  my PhD. They were the yin and yang in my PhD, and the body of work
  you read in this thesis was a direct result of this
  delicate balance. Max always ensured that my ideas were grounded and
  would work on a robotic system, which was especially useful as I often
  flew away with idealistic ideas. I admire his extreme patience
  during our meetings and his approach of understanding complex ideas
  from fundamentals. If not these amazing skills, I hope I learned how
  to always stay humble and listen to collaborators from him. Drew has been a
  constant source of inspiration throughout my journey and allowed me
  to chase crazy ideas purely because they were interesting. For someone
  who is a CTO of a self-driving company, his passion for pure fundamental
  research is very inspiring and has driven me to work harder on ideas
  that I was excited about. I loved all our phone calls, with me
  speaking cryptic equations while he somehow miraculously deciphers
  them, our walks on campus discussing ideas and papers, and our terse
  emails with the entire content in the subject and no body. The most
  fun projects in my PhD were a result of brainstorming with Drew, and
  for that I am extremely grateful.

  % Committee - Oliver, Leslie
  I take this opportunity to also thank my committee members,
  Oliver and Leslie. I interacted with Oliver a lot of times during my
  PhD both as an RI student who was interested in his work, and as his
  teaching assistant for two courses. I admire his straightforward way
  of thinking and insights in to the robot manipulation problem, which
  influenced my own research. Leslie has always been one of my idols
  and for someone entering into the field of robot planning and
  learning, some of her works were foundational. I am extremely lucky
  to have them both on my committee and thankful for all the feedback
  I have received from them on my thesis.

  % Apple - Martin, Chris, Humphrey, Jakub
  My journey also took me to some wonderful places besides CMU. I was
  fortunate to work with Martin Levihn, Chris Clark, Humphrey Hu, and
  Jakub Dworakowski at Apple Special Projects Group. All of them have
  been incredible mentors both professionally and personally. Their
  insights and guidance has influenced both this thesis and my future
  beyond PhD. I also worked with Vladlen Koltun and Ozan Sener at
  Intel Labs, where I learned how to tackle extremely difficult
  problems and the value of collaboration. I'd like to thank Ozan for
  his patience while I slowly grasped the
  challenges of the problem.

  % Masters Advisors - Jean, Katharina
  I would be remiss if I did not thank Jean Oh and Katharina Muelling
  who were my mentors at the start of this journey. They showed me how
  exciting research could be, and encouraged me to apply for a
  PhD. Jean is one of the nicest people I know, and I am
  fortunate to have been her student.

  % SBPL - Dhruv, Suhail, Fahad, Sung, Andrew, Tushar, Yash, Venkat, Jacky,
  % Anahita, Shivam, Ram, Oren, Vinitha
  During my time at CMU, I was a part of two wonderful labs - SBPL and
  Lairlab. I would like to thank my officemates, Dhruv Saxena, Yash Oza, and
  Tushar Kusnur for putting up with my constant scribbles on the whiteboard
  and my loud discussions. I would like to thank Dhruv especially for
  always lending an ear when I needed it. For the first 3 years of my
  PhD, I was part of a truck unloading project and worked with fellow
  SBPL members - Oren
  Salzman, Fahad Islam,
  Sung-Kyun Kim, and Andrew Dornbush, who were all insanely helpful whenever I got stuck
  and helped me learn motion planning and belief space planning
  fundamentals. Besides work, I had loads of fun with other SBPL
  friends - Venkatraman Narayanan, Jacky Liang, Anahita Mohseni Kabir,
  Shivam Vats, Ram Natarajan, Manash Pratim Das, Vinitha Ranganeni,
  Shohin Mukherjee, and 
  Rishi Veerapaneni. While SBPL was where I physically most of my time, Lairlab
  was my go to for brainstorming on machine learning research. Wen Sun
  was my close collaborator initially in my PhD, and he almost became
  my third advisor for a year. Fellow ``Bagnellians'' -  Allie Del
  Giorno, Hanzhang Hu, Jiaji 
  Zhou, Wen Sun, Arun Venkatraman, and Gokul Swamy
  were wonderful labmates and I will always cherish the lab lunches
  with them which usually had heated discussions on topics ranging
  from research to the latest gossip. Special shout out to Allie for
  making my defense super fun, and helping me with reviewing all my
  papers and presentations. 

  % Lairlab - Allie, Wen, Gokul, Echo, Jiaji, Arun

  % RI friends - Achal, Xuning, Alex, Adithya, Jerry, Puneet, Vishal, Senthil,
  % Ankit, Rosario, Rogerio, Arjav, Eric, Tabitha, Saumya, Dominik, Cornelia,
  % Yifan, Debi, Nick, Abhijeet, Sanjiban, Ratnesh
  The robotics institute at CMU has a wonderful student community. As
  a part of RoboOrg, and as a RI Masters and PhD student, I had the good
  fortune to know some of the most amazing and talented
  individuals. These friendships lasted over 6 years and will last for
  a lot more years to come. I would like to thank
  Puneet Puri, Vishal Dugar, Jerry Hsiung, Senthil Purushwalkam, Ankit Bhatia,
  Adithya Murali, Alex Spitzer, Xuning Yang, Achal Dave, Debidatta
  Dwibedi, Rogerio
  Bonatti, Zhi Tan, Rosario Scalise, Eric Huang, Arjav Desai, Tabitha Lee,
  Yifan Hou, Brian Okorn, Tanmay Shankar, Roberto Shu, Sibi
  Venkatesan, Nick
  Gisolfi, Abhijeet Tallavajhula, Ratnesh Madaan, Sankalp Arora, and
  Sanjiban Choudhury. A special shout-out to NSH 2204 for being
  legendary, you know you are. I would like to give special thanks to
  Sanjiban who was my first mentor at CMU, and remained a friend
  I always reach out to for guidance on work, life, and beyond.

  % RI admin - Suzanne, Jean
  Suzanne Muth and Jean Harpley made my life at RI feel seamless. Suzanne
  worked hard behind the scenes to ensure that RI PhDs could focus
  solely on research. Jean Harpley is one of the most amazing people
  in RI. She helped RoboOrg and the student community in RI at every
  step, even though it was not a part of her job description. For that
  and many more, I would like to thank both of them.

  % CMU friends - Abilash, Ben, Shefali, Nikita, Brandon, Cara, Ananya
  RI was too small to contain all the wonderful people I know, and I
  met people from other departments who played a big part in my
  journey. Abilash Subbaraman was always down to climb with me and do
  the craziest, and often the most dangerous, adventure sports with
  me. Ben Eysenbach helped me understand how my research fits with the
  broader RL literature. Shefali Umrania and Ananya Uppal were close
  friends who made my grad school life extremely fun. The Explorers
  club of Pittsburgh introduced me to several inspiring and fun
  people with whom I went backpacking, climbing, and biking
  with. All work and no play could have made this journey dull, but
  these folks ensured it was anything but dull.

  % Family - Mom, Dad, Venkat, Payal
  I am the person I am today because of my family. My dad instilled
  the value of working hard in me, and pushed me to always have bigger
  ambitions. My brother is my constant partner-in-crime and was a huge
  inspiration in my career choices. He has always prioritized my
  choices over his own, and baked a fear of mediocrity in me, that
  has always pushed me to go further. My mom is the reason for all my
  successes, and her countless sacrifices gave me the opportunity to
  focus on my education while she took care of far more important
  issues. I hope I say it enough, but if I don't here it goes - I am
  extremely grateful to be your son, brother, and a part of this
  family. 

  % Love - Pragna
  I said PhD is all about the journey. This journey started with me
  meeting my partner Pragna Mannam, and she enriched it far beyond
  what it would have been without her. She always believed in me, when
  I did not. She brought me to the surface, when I sank to the
  depths. She supported me unconditionally, when I needed it. She is
  my rock, and the shining star in my sky. At the end of this journey,
  I am at a loss of words to thank her, and can only say ``Cutie, we
  did it!''
\end{acknowledgments}
\restoregeometry

\pagestyle{plain}

\tableofcontents
\addtocontents{toc}{\vspace*{-2cm}}
\listoffigures
\addtocontents{lof}{\vspace*{-2cm}}
\listoftables
\listofalgorithms

\mainmatter

\input{chapters/introduction}
\input{chapters/background}
%\input{chapters/related-work}
\input{chapters/aistats19}
\input{chapters/cmax}
\input{chapters/cmaxpp}
\input{chapters/ilc}
%\input{chapters/proposed-work}
\input{chapters/taml}
\input{chapters/future-work}
\input{chapters/appendix}

\chapter*{Bibliography}
\addcontentsline{toc}{chapter}{Bibliography}

\vspace{-25mm}
This bibliography contains \total{citenum} references.
\vspace{10mm}

\printbibliography[heading=none]

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
